---
title: "Collatz Messing About - Breaking some rules"
author: "Phileas Dazeley Gaist"
date: "11/29/2021"
output: 
  html_document:
    df_print: paged
  md_document:
    variant: markdown_github
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

# Breaking some rules

Let's break some rules and see if we can learn anything! What changes when we change 
the circumstances in which we compute Collatz - or similar - sequences?

Let's relax and mess about!

## Sequence generating function

Firstly, let's diverge from the rules by making a custom sequence-generation 
function, which allows us to change the rules for even and odd numbers, and specify 
a start value, goal value, and maximum number of steps:

```{r}
sequence_even <- function(even_value, factor_value, added_value){
  return(factor_value * even_int + added_value)
}

sequence_odd <- function(odd_value, factor_value, added_value){
  return(factor_value * odd_int + added_value)
}

sequence_generator <- function(start_value, goal_value = 1, 
                               even_fact = 0.5, even_add = 0, 
                               odd_fact = 3, odd_add = 1,
                               max_steps = 1000){
  new_val <- start_value
  steps <- c(new_val)
  step_count <- 0
  
  while(new_val != 1) {
    step_count <- step_count + 1
    if(new_val %% 2 == 0){
      new_val <- sequence_even(new_val, even_fact, even_add)
      steps <- c(steps, new_val)}
    else{
      new_val <- sequence_odd(new_val, odd_fact, odd_add)
      steps <- c(steps, new_val)}
    if(step_count > max_steps){
      break
    }
  }
  return(list(steps))
}
```

Now we can make the rules vary across 4 dimensions: the factors in either rule, 
and the additions in either rule. 

Let's see it in action! 

## Alternate sequence types

First, let's create a few vectors of parameters to enter into our function. For 
factor values, I chose ten below 1, and 10 above. Factor values below 1 are -1, and 
values from 0.1 to 0.9 separated by 0.1 increments, while values above are 2 to 20. 
For addition values, I chose values from 1 to 20, separated by single integer 
increments.

Why these values? It's hard to say exactly but my instinct was that it might be 
interesting that way: Reversing the vectors for left and right makes it easy to 
keep the sequence generation parameters different for left and right, having the 
same number of parameters is convenient if I want to use some types of loops, and 
I liked the idea of having half of the factor values below 1, and the other half above.

Of course, I'd like to also test sequence generations with the same odd and even 
rules (which will likely/almost surely be a lot more/completely generally predictable), 
and sequence generations with randomised parameters from these vectors, and perhaps 
others too! 

```{r}
even_factors <- c(-1, seq(0.1, 0.9, by = 0.1), seq(1, 10, by = 1))
even_additions <- c(seq(1, 20, by = 1))
odd_factors <- rev(even_factors)
odd_addition <- rev(even_additions)
```

Now we need to collect some data:

```{r}
seeds <- 1:10
max_s <- 1000
```

We can collect and group the data in data frames by sequence varieties or by 
starting values.

By sequence varieties:

```{r}

```

By starting values: 

```{r}

```

Ideas of things to try:

- taking random values in a range and plotting their trajectories
- plotting the trajectories of decimal values
- plotting the trajectories of even and odd numbers in different colours
- computing and plotting trajectories for different rules
- computing the trajectories of negative numbers
- computing the trajectories of imaginary and complex numbers (possibly with a 
modification for complex numbers which are neither even nor odd, or by using modified
rules which would force the values to be either odd or even)
- look at patterns in the data, run statistical tests on them, and maybe forms of 
correlation tests and regressions
- plots of sequences with different parameters on top of each other to see how the 
parameters make the end sequences vary (randomised or ordered, or the same values)
- plot a visualisation of the latent space between sequences of different parameters
- make animations between parameters and save them as GIF images